Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics and Predicts Long Timescale Behavior in Nanoscale Pores.

### 2019-12-03 Ben_Manuscripts/stochastic_transport/Draft.tex
New ways to fit FPT distribution

  \begin{abstract}
  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories, in order
  to predict solute mean squared displacements (MSDs). With our first model, based
  on the anomalous diffusion literature, we show how solute dynamics can be modeled as
  a fractional diffusion process subordinate to a continuous time random walk.
  We parameterize the distribution of dwell times, hop lengths between dwells and
  correlation between hops. We explore 2 variations of the model. The first variation
  applies a single set of parameters to the solutes and the second applies 2 sets
  of parameters based on the solute's radial distance from the closest pore center.
  Our second model, an adaptation of Markov state models, treats the configurational
  state of the system as a Markov process, each of which has state dependent 
  transport behavior. For each state and transition between states, we parameterized
  the distribution and temporal correlation structure of positional fluctuations as
  a means of characterization and to allow us to predict solute MSDs.
  We show that our models reasonably reproduce MSDs calculated from molecular 
  simulations. The anomalous diffusion model shows superior predictive performance, 
  in many cases reproducing MD results well within error. Finally, we use the 
  parameters of each model to identify specific differences in solute transport
  behavior which can help suggest modifications that can be made to LLC membrane 
  monomers in order to increase or reduce selectivity towards a specific solute
  or class of solutes.

@@

@@

  %BJC4: another way to visualize data
  %BJC4: I think these are kind of hard to read. Can't really see any trends.
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{1mode_radar.pdf}
  \caption{}\label{fig:1mode_radar}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{2mode_radar.pdf}
  \caption{}\label{fig:2mode_radar}
  \end{subfigure}
  \caption{Another way to visualize Tables~\ref{table:sfbm_params} and~\ref{table:sfbm_params_2mode}}\label{fig:radars}
  \end{figure}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{1mode_radar_nolines.pdf}
  \caption{}\label{fig:1mode_radar}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{2mode_radar_nolines.eps}   % for some reason it has trouble saving this as a PDF (.eps doesn't show opacity)
  \caption{}\label{fig:2mode_radar}
  \end{subfigure}
  \caption{Same as Figure~\ref{fig:radars} but with no lines between points}\label{fig:radars_nolines}
  \end{figure}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_hops.pdf}
  \caption{$\mathcal{N}(\sigma) / L(\sigma, \alpha_h)$}\label{fig:1mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_dwells.pdf}
  \caption{$P(\alpha_d) / P_T(\alpha_d, \lambda)$}\label{fig:1mode_AD_dwells}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hurst_barchart.pdf}
  \caption{$\gamma(H)$}\label{fig:hurst_barchart}
  \end{subfigure}
  \caption{The parameters of the 1 mode model reveal differences in solute
	  dynamics. (a) We parameterized Gaussian ($\mathcal{N}(\sigma)$,
	  parameters given by bars with white background) and L\'evy stable
	  ($L(\sigma, \alpha_h)$, parameters given by bars with gray
	  background) distributions to describe solute hop lengths. We assume
	  the mean ($\mu$) to be zero for these distributions and there to be
	  no skewness ($\beta=0$) in the L\'evy stable distributions. High
	  values of $\sigma$ and lower values of $\alpha_h$ result in larger
	  hops. (b) We parameterized a pure power law ($P(\alpha_d)$, white
	  background) and a truncated power law ($P_T(\alpha_d, \lambda)$, gray
	  background) distribution to describe solute dwell times. Lower values
	  of $\alpha_d$ lead to heavier power law tails and higher values of
	  $\lambda$ truncate the distribution at lower dwell times.(c) Finally,
	  we parameterized the hop autocorrelation function ($\gamma(H)$) to
	  describe the degree of correlation between hops.  Higher values of
	  $H$ display closer to Brownian behavior.
	  }\label{fig:1mode_bar_charts}
  \end{figure}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{2mode_AD_hops.pdf}
  \caption{$\mathcal{N}(\sigma) / L(\sigma, \alpha_h)$}\label{fig:2mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{2mode_AD_dwells.pdf}
  \caption{$P(\alpha_d) / P_T(\alpha_d, \lambda)$}\label{fig:2mode_AD_dwells}
  \end{subfigure}
  \caption{The two mode model parameterizes solute behavior in the pores and
	  tails separately. Mode 1 corresponds to solute behavior within 0.75
	  nm of the pore center, the pore region, and mode 2 corresponds to
	  behavior greater than or equal to 0.75 nm from the pore center, the
	  tail region. (a) Generally, movement is much more restricted in the
	  tail region shown by lower mode 2 $\sigma$ values for the Normal
	  (white background) and L\'evy distributions (gray background). Values
	  of $\alpha$ are significantly lower for urea and acetic acid meaning
	  there is a larger probability that they will take large hops.
	  (b) Dwell times are longer in in the tails. Lower values of
	  $\alpha_d$ correspond to power laws with heavier tails and thus
	  higher probabilities of long dwell times. There is no easily
	  discernable trend in $\lambda$ of the truncated power law
	  distribution (gray background).  Note that we used the same Hurst
	  parameter for both modes due to a low number of sufficiently long
	  sequences of hops in each mode. See Figure~\ref{fig:1mode_bar_charts}
	  for descriptions of the parameters.
  }\label{fig:2mode_bar_charts}
  \end{figure}

>>>>

  \begin{abstract}
  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories, in
  order to predict solute mean squared displacements (MSDs) and mean first 
  passage times (MFPTs) in macroscopic length pores. With our first model, 
  based on the anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore 2 variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies 2 sets of parameters based on the solute's radial distance from the closest pore center.
  Our second model, a generalization of Markov state models, treats the configurational
  state of the system as a Markov process, each of which has state dependent 
  transport behavior. For each state and transition between states, we parameterize
  the distribution and temporal correlation structure of positional fluctuations as
  a means of characterization and to allow us to predict solute MSDs.
  We show that our models reasonably reproduce MSDs calculated from molecular 
  simulations. The anomalous diffusion model shows superior predictive performance, 
  in many cases reproducing MD results well within error. Finally, we demonstrate 
  how one can use our models in order to calculate mean first passage time, and 
  therefore solute flux, of a given solute across a macroscopic-length pore. Overall,
  this works helps connect the relationship between microscopic chemically-dependent
  solute motions and macroscopic membrane performance.

%  Finally, we use the 
%  parameters of each model to identify specific differences in solute transport
%  behavior which can help suggest modifications that can be made to LLC membrane 
%  monomers in order to increase or reduce selectivity towards a specific solute
%  or class of solutes.

@@

  \begin{equation}
  F(t) = erfc\bigg(\frac{x - vt}{2\sqrt{D t}}\bigg)
  \end{equation}

@@

  %BJC4: another way to visualize data
  %BJC4: I think these are kind of hard to read. Can't really see any trends. Open to other
  %ideas.
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{1mode_radar.pdf}
  \caption{}\label{fig:1mode_radar}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{2mode_radar.pdf}
  \caption{}\label{fig:2mode_radar}
  \end{subfigure}
  \caption{Another way to visualize Tables~\ref{table:sfbm_params} and~\ref{table:sfbm_params_2mode}}\label{fig:radars}
  \end{figure}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{1mode_radar_nolines.pdf}
  \caption{}\label{fig:1mode_radar}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\linewidth]{2mode_radar_nolines.png}   % for some reason it has trouble saving this as a PDF
  \caption{}\label{fig:2mode_radar}
  \end{subfigure}
  \caption{Same as Figure~\ref{fig:radars} but with no lines between points}\label{fig:radars_nolines}
  \end{figure}

@@
